### 2021-03-31 header.tex
Initial Overleaf Import

%header

%ABNT Packages
\usepackage[utf8]{inputenc}
\usepackage[lmargin=3cm,tmargin=3cm,rmargin=2cm,bmargin=2cm]{geometry}
\usepackage[onehalfspacing]{setspace}
\usepackage[T1]{fontenc}
%\usepackage[brazil]{babel}

% Keywords command
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

%Essential Packages
\usepackage{graphicx,xcolor,comment,enumerate,multirow,multicol,indentfirst}

%Math Packages
\usepackage{amsmath,amsthm,amsfonts,amssymb,dsfont,mathtools,blindtext}

\usepackage{soul}
\usepackage{textcomp} 
\usepackage{gensymb}
\usepackage{float}
\usepackage{lineno,hyperref}
\usepackage{booktabs} 
\usepackage{calc}
\usepackage{caption}