### 2021-04-01 abstract.tex
visual computer version

\begin{abstract}
	\textbf{Image analysis is an essential field for several topics of life sciences, such as biology or botany. In particular, seeds analysis (e.g., fossil research) can provide significant information about their evolution, the history of agriculture, the domestication of plants, and the knowledge of diets in ancient times. This work aims to present a software that performs an image analysis by feature extraction and classification starting from images containing seeds through a brand new and unique framework. In detail, we propose two \emph{ImageJ} plugins, one capable of extracting morphological, textural, and colour characteristics from images of seeds, and another one to classify the seeds into categories by using the extracted features. The experimental results demonstrated the correctness and validity both of the extracted features and the classification predictions. The proposed tool is easily extendable to other fields of image analysis.} 
\end{abstract}