### 2021-04-01 header.tex

%Essential Packages
\usepackage{graphicx,xcolor,comment,enumerate,multirow,multicol,indentfirst}

\usepackage{soul}
\usepackage{textcomp} 
\usepackage{gensymb}
\usepackage{float}
\usepackage{lineno,hyperref}
\usepackage{booktabs} 
\usepackage{calc}
\usepackage{caption}

\usepackage{authblk}

\smartqed  % flush right qed marks, e.g. at end of proof

\RequirePackage{fix-cm}
\usepackage{natbib}