### 2021-04-01 main.tex

\documentclass{svjour3}                     % onecolumn (standard format)

\smartqed  % flush right qed marks, e.g. at end of proof

\input{header}

\journalname{The Visual Computing}

\begin{document}

	\title{An effective and friendly tool for seed image analysis}

	%\titlerunning{Short form of title}        % if too long for running head

	\author{Loddo A. \and
		Di Ruberto C. \and
		Vale A.M.P.G. \and
		Ucchesu M. \and
		Soares J.M. \and
		Bacchetta G.
	}

	%\authorrunning{Short form of author list} % if too long for running head

	\institute{Loddo A., Di Ruberto C.\at
		Department of Mathematics and Computer Science, University of Cagliari, \\
		via Ospedale 72, 09124 Cagliari, Italy. \\
		\email{andrea.loddo@unica.it}
		\and
		Vale A.M.P.G.\at
		Escola Agrícola de Jundiaí (EAJ), Universidade Federal do Rio Grande do Norte (UFRN), \\
		Rodovia RN 160, Km 03, CEP 59280-000 Macaíba (RN), Brasil.
		\and
		Ucchesu M., Bacchetta G. \at
		Centro Conservazione Biodiversità (CCB), Dipartimento di Scienze della Vita e dell’Ambiente (DiSVA), Università degli Studi di Cagliari, \\
		Viale S. Ignazio da Laconi 13, 09123 Cagliari, Italy.\\
		\and
		Soares J.M. \at
		Universidade Federal do Rio Grande do Norte (UFRN), \\
		Rodovia RN 160, Km 03, CEP 59280-000 Macaíba (RN), Brasil.\\
		\and
		Bacchetta G. \at
		Hortus Botanicus Karalitanus (HBK), Università degli Studi di Cagliari, \\
		Viale S. Ignazio da Laconi 11, 09123 Cagliari, Italy.\\		
	}

	\date{Received: date / Accepted: date}
	% The correct dates will be entered by the editor

	\maketitle

	\input{abstract.tex}

	\keywords{image analysis, features, classification, biology, seeds, imagej}

	\input{intro.tex}

	\input{material.tex}

	\input{plugin.tex}

	\input{results.tex}

	\input{conclusion.tex}

	%\begin{acknowledgements}
	%If you'd like to thank anyone, place your comments here
	%and remove the percent signs.
	%\end{acknowledgements}

	\section*{Conflict of interest}
	The authors declare that they have no conflict of interest.

	% BibTeX users please use one of
	\bibliographystyle{spbasic}      % basic style, author-year citations
	\bibliography{bib}   % name your BibTeX data base

\end{document}