### 2021-07-12 extra/conflicts.tex

\documentclass{letter}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-0.5in}
\setlength{\oddsidemargin}{0pt}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false,linkcolor=blue,citecolor=blue]{hyperref}

\address{Andrea Loddo\\
	Dept. of Mathematics and Computer Science\\
	University of Cagliari \\
	09128, Cagliari, Italy \\
	Phone:(+39) 070 675 8503 \\
	Email: \texttt{andrea.loddo@unica.it}}
\signature{Andrea Loddo}

\include{macro}

\begin{document}
	\begin{letter}{}

		\opening{Dear Editors and Reviewers}

		The authors declare no conflict of interest.

		\closing{Yours sincerely,}
	\end{letter}

\end{document}